---
title: "R Practice"
output: 
  html_document:
    toc: true # table of content true
    toc_depth: 3  # upto three depths of headings (specified by #, ## and ###)
    toc_float:
      collapsed: false
      smooth_scroll: false
    df_print: paged
  # 
  # number_sections: true  ## if you want number sections at each table header
  # theme: united  # many options for theme, this one is my favorite.
  # highlight: tango  # specifies the syntax highlighting style
---

```{r setup, include=F}
knitr::opts_chunk$set(echo = TRUE)
```


```{r class.source="bg-danger", class.output="bg-warning"}
#setwd("type/here/the/address/")
setwd("C:/Users/soumy/OneDrive/Desktop/assignments/s and p")
```


#### just to make sure that the file is there let's see which files are in this directory<br>

```{r class.source="bg-danger", class.output="bg-warning"}
#setwd("C:/Users/soumy/OneDrive/Desktop/assignments/s and p")
list.files()
```

# Problem Set R

### Now is yoiur turn to practice!!

#### We are using a new Dataset, This is a frequently used dataset for multiple applications in statistics and machine learning. This dataset is deposited in a website that is pretty useful in bioinformatics to look for tutorials, datasets and advice Kaggle (https://www.kaggle.com)

#### Go to https://www.kaggle.com/saurabh00007/diabetescsv and download the csv file

#### This dataset consists on clinical variables for 768 patients to evaluate a few variables to predict whether a patient has diabetes.
library(dplyr)

#### Please write the R code necessary to run the next items:

#### 1. Load the dataset and show the first 5 lines<br>
```{r}
my_data <- read.csv("diabetes.csv")
head(my_data, 5)
```

#### 2. How many patients have diabetes?<br>
```{r}
num_diabetes <- sum(my_data$Outcome == 1)
print(paste("Number of patients with diabetes:", num_diabetes))
```

#### 3. How many patients have diabetes that are older than 45?<br>
```{r}
num_diabetes_over_45 <- sum(my_data$Outcome == 1 & my_data$Age > 45)
print(paste("Number of patients with diabetes older than 45:", num_diabetes_over_45))
```
#### 4. What is the mean and variance of glucose levels for individuals without diabetes<br>
```{r}
mean_glucose_no_diabetes <- mean(my_data$Glucose[my_data$Outcome == 0], na.rm = TRUE)
var_glucose_no_diabetes <- var(my_data$Glucose[my_data$Outcome == 0], na.rm = TRUE)
print(paste("Mean glucose (no diabetes):", mean_glucose_no_diabetes))
print(paste("Variance of glucose (no diabetes):", var_glucose_no_diabetes))
```

#### 5. Create a new discrete variable that has 1 if the individual has diabetes and high blood pressure (above 100), 2 if an indivual has diabetes and low blood pressure and 3 if the individual does not have diabetes.<br>
```{r}
my_data$Diabetes_BP_Category <- ifelse(my_data$Outcome == 1 & my_data$BloodPressure > 100, 1, 
                             ifelse(my_data$Outcome == 1 & my_data$BloodPressure <= 100, 2, 
                                    ifelse(my_data$Outcome == 0, 3, NA)))

table(my_data$Diabetes_BP_Category)
```

#### 6. Construct two plots of the distribution of BMI for individuals with diabetes and without diabetes<br><br>

```{r}
library(ggplot2)

# Filtering the data for individuals with and without diabetes
diabetes_data <- my_data[my_data$Outcome == 1, ]
no_diabetes_data <- my_data[my_data$Outcome == 0, ]

par(mfrow=c(1, 2))  # Set up a 1x2 grid for side-by-side plots

# Plot for individuals with diabetes
hist(diabetes_data$BMI, 
     main="BMI Distribution for Individuals with Diabetes", 
     xlab="BMI", 
     col="red", 
     border="black", 
     breaks=20)

# Plot for individuals without diabetes
hist(no_diabetes_data$BMI, 
     main="BMI Distribution for Individuals without Diabetes", 
     xlab="BMI", 
     col="blue", 
     border="black", 
     breaks=20)
```


<hr class="rounded">

